Tight Generalization Bounds for Hard Parameter Sharing in High-dimensional Linear Regression

### 2020-10-10 arxiv/proof_same_cov.tex
sec A

	\paragraph{Part 3:} The last part is the square of $E_j$,
	\begin{align}\label{eq_gA_p3}
		\sum_{j=1}^t \norm{E_j}^2 &= \bignormFro{\cW A - \cE}^2
		= \bignormFro{\cE}^2 - \inner{\cW A}{\cE},
	\end{align}
	which is because $\norm{\cW A}^2 = \inner{\cW A}{\cE}$ by algebraic calculation.
	Hence, it suffices to show that the expectation of equation \eqref{eq_gA_p3} is equal to $\sigma^2 (n\cdot t - p\cdot r)$.
	First, we have that $\ex{\|\cE\|_F^2} = \sigma^2 \cdot n \cdot t$.
%	Conditional on $X$, we have that
%	\begin{align*}
%		\exarg{\cE}{\norm{E_j}^2}
%		= \exarg{\cE}{\sum_{j=1}^t \bigbrace{\bignorm{\cW A_j}^2 - 2\inner{\cW A_j}{\varepsilon^{(j)}} }} + \sigma^2 \cdot n \cdot t.
%	\end{align*}
%	We consider both terms in the above equation one by one.
	Second, notice that
	%\begin{align}
	%	\sum_{j=1}^t \norm{\cW A_j}^2 = \sum_{j=1}^t \bigtr{\cW A_j A_j^{\top} \cW^{\top}} = \bigtr{\cW AA^{\top} \cW^{\top}} = \bigtr{U_X U_X^{\top} \cE A^{\top} (AA^{\top})^{-1} A \cE^{\top}}. \label{eq_proof_same_cov_1}
	%\end{align}
	%We observe that
	\begin{align*}
			\exarg{\cE}{\cE A^{\top} (AA^{\top})^{+} A \cE^{\top}}
		&= \exarg{\cE}{\sum_{j=1}^t \sum_{k=1}^t \varepsilon^{(j)} A_j^{\top} (AA^{\top})^{+} A_k {\varepsilon^{(k)}}^{\top}} \\
		&= \exarg{\cE}{\sum_{j=1}^t \varepsilon^{(j)} A_j^{\top} (AA^{\top})^{+} A_j {\varepsilon^{(j)}}^{\top}} \\
		&= \sigma^2 \cdot r \cdot \id_{n\times n}.
	\end{align*}
	In the above derivation, the second step used the fact that for any $j\neq k$, $\varepsilon^{(j)}$ and $\varepsilon^{(k)}$ are pairwise independent.
	The third step used the fact that $\sum_{j=1}^t A_j^{\top} (AA^{\top})^{+} A_j = \bigtr{\id_{r\times r}} = r$, and $\ex{\varepsilon^{(j)} {\varepsilon^{(j)}}^{\top}} = \sigma^2 \cdot \id_{n\times n}$.
	Therefore, we have that
%	\begin{align*}
%		\inner{\cW A}{\cE} = \sigma^2 \cdot r \cdot \bigtr{U_X U_X^{\top}} = \sigma^2 \cdot r \cdot p,
%	\end{align*}
%because $U_X$ has rank $p$ by Fact \ref{lem_minv}.
	\begin{align*}
		\inner{\cW A}{\cE} = \sigma^2 \cdot r \cdot \bigtr{X(X^{\top} X)^{-1} X^{\top}} =  \sigma^2 \cdot r \cdot \bigtr{ X^{\top}X(X^{\top} X)^{-1}}  = \sigma^2 \cdot r \cdot p.
	\end{align*}
   because $X^\top X$ is a $p\times p$ matrix.	
	%Next, note that
	%\[ \sum_{j=1}^t \inner{\cW A_j}{\varepsilon^{(j)}} = \bigtr{\cW A\cE^{\top}}, \]
	%which is equal to equation \eqref{eq_proof_same_cov_1} by the definition of $\cW$.
	Hence the proof is complete.

@@

	For equation \eqref{eq_gA_p3}, using Lemma \ref{largedeviation} and the fact that all moments of $\varepsilon_i$ exist, we obtain that with high probability
	\begin{align}
		\abs{\normFro{\cE}^2 - \ex{\normFro{\cE}^2}} \le p^{-c_{\infty}} \ex{\normFro{\cE}^2}. \label{eq_gA_err3}
	\end{align}
	For the inner product between $\cW A$ and $\cE$, we have that
	\begin{align}
		\bigabs{\inner{\cW A}{\cE} - \ex{\inner{\cW A}{\cE}}} &= \bigabs{\bigtr{U_X U_X^{\top} (\cE U_AU_A^{\top} \cE^{\top} - \sigma^2 \cdot r \id_{n\times n})}} \nonumber  \\
		&\le \bignormFro{U_{X} U_X^{\top}} \cdot \bignorm{\cE U_A U_A^{\top} \cE^{\top} - \sigma^2 \cdot r \id_{n\times n}} \nonumber \\
		&\le p^{1/2} \cdot (\sigma^2 r\cdot p^{\e}). \label{eq_gA_err4}
	\end{align}
	%\begin{align*}
	%	%|\inner{U_X U_X^{\top} \bigbrace{\varepsilon_i W_i^{\top}} (\cW\cW^{\top})^{-1}W_j}{\varepsilon_j}|
	%	\abs{W_i^{\top} (\cW\cW^{\top})^{-1} W_j \cdot \bigbrace{\varepsilon_i^{\top} U_X U_X^{\top} \varepsilon_j}}
	%	&\le |W_i^{\top} (\cW\cW^{\top})^{-1}W_j| \cdot \sigma^2 p^{\e} \bigtr{(U_X U_X^{\top})^2}^{1/2} \le \sigma^2 p^{1/2+\e}
  % &\left|\left(1-\mathop{\mathbb{E}}_{\varepsilon_j} \right)\left[\inner{U_X U_X^{\top} \bigbrace{\varepsilon_j W_j^{\top}} (\cW\cW^{\top})^{-1}W_j}{\varepsilon_j}\right]\right| \le |W_j^{\top} (\cW\cW^{\top})^{-1}W_j| \cdot \sigma^2 p^{\e} \tr \left[(U_X U_X^{\top})^2\right]\le \sigma^2 p^{1/2+\e},\\
  % &\left|\left(1-\mathop{\mathbb{E}}_{\varepsilon_j} \right)\left[\inner{U_X U_X^{\top} \bigbrace{\varepsilon_j W_j^{\top}} (\cW\cW^{\top})^{-1}W_j}{U_X U_X^{\top} \bigbrace{\varepsilon_j W_j^{\top}} (\cW\cW^{\top})^{-1}W_j}\right]\right|\\
  % &\le |W_j^{\top} (\cW\cW^{\top})^{-1}W_j|^2 \cdot \sigma^2 p^{\e} \tr \left[(U_X U_X^{\top})^4\right]\le \sigma^2 p^{1/2+\e},
	%\end{align*}
	%Combining equation \eqref{eq_gA_}, we get that with high probability
	%\begin{align}
	%	\left|g_2(\cal W)-\sigma^2(n\cdot t - p\cdot r)\right|=\left|g_2(\cal W)- \exarg{\varepsilon_1, \dots, \varepsilon_t} {g_2(\cal W)}\right| \lesssim \sigma^2 p^{1/2+\e}. \label{eq_gA_err}
	%\end{align}
	Combining equation \eqref{eq_gA_err1}, \eqref{eq_gA_err2}, \eqref{eq_gA_err3}, and \eqref{eq_gA_err4}, we obtain that equation \eqref{eq_gA_err} holds.

>>>>

	\paragraph{Part 3:} The last part is the square of $E_j$,
	\begin{align}\label{eq_gA_p3}
		\sum_{j=1}^t \norm{E_j}^2 &= \bignormFro{\cW A - \cE}^2
		= \bignormFro{\cE}^2 - \inner{\cW A}{\cE},
	\end{align}
	which is because $\norm{\cW A}^2 = \inner{\cW A}{\cE}$ by algebraic calculation.
	Hence, it suffices to show that the expectation of equation \eqref{eq_gA_p3} is equal to $\sigma^2 (n\cdot t - p\cdot r)$.
	First, we have that $\ex{\|\cE\|_F^2} = \sigma^2 \cdot n \cdot t$.
%	Conditional on $X$, we have that
%	\begin{align*}
%		\exarg{\cE}{\norm{E_j}^2}
%		= \exarg{\cE}{\sum_{j=1}^t \bigbrace{\bignorm{\cW A_j}^2 - 2\inner{\cW A_j}{\varepsilon^{(j)}} }} + \sigma^2 \cdot n \cdot t.
%	\end{align*}
%	We consider both terms in the above equation one by one.
	Second, notice that for $U_XU_X^\top:=X(X^{\top} X)^{-1} X^{\top}$ and $1\le i, j \le t$,
	%\begin{align}
	%	\sum_{j=1}^t \norm{\cW A_j}^2 = \sum_{j=1}^t \bigtr{\cW A_j A_j^{\top} \cW^{\top}} = \bigtr{\cW AA^{\top} \cW^{\top}} = \bigtr{U_X U_X^{\top} \cE A^{\top} (AA^{\top})^{-1} A \cE^{\top}}. \label{eq_proof_same_cov_1}
	%\end{align}
	%We observe that
	\begin{align*}
			\exarg{\cE}{(\cE^{\top}U_XU_X^\top \cE)_{ij} }
		&= \exarg{\cE}{{\varepsilon^{(i)}}^\top U_X U_X^\top  \varepsilon^{(j)} } = \sigma^2 \cdot \tr\left( U_XU_X^\top \right) \cdot \delta_{ij} = p\sigma^2 \cdot \delta_{ij},
	\end{align*}
	which implies that $\exarg{\cE}{\cE^{\top}U_XU_X^\top \cE }=p\sigma^2 \cdot \id_{t\times t}$.
	In the above derivation, the second step used the fact that $\E(\varepsilon^{(i)}_k \varepsilon^{(i)}_l )=\sigma^2\cdot \delta_{ij}\delta_{kl}$ for $1\le k,l \le n$, and the third step used the fact that $\tr(U_XU_X^\top) =\tr(U_X^\top U_X) =p$.
	Therefore, we have that
%	\begin{align*}
%		\inner{\cW A}{\cE} = \sigma^2 \cdot r \cdot \bigtr{U_X U_X^{\top}} = \sigma^2 \cdot r \cdot p,
%	\end{align*}
%because $U_X$ has rank $p$ by Fact \ref{lem_minv}.
	\begin{align*}
		\exarg{\cE}{ \inner{\cW A}{\cE}} = p\sigma^2 \cdot \bigtr{\id_{t\times t}U_AU_A^\top} =  p \sigma^2 \cdot r ,
			\end{align*}
   because $U_AU_A^\top$ has rank $r$.	
	%Next, note that
	%\[ \sum_{j=1}^t \inner{\cW A_j}{\varepsilon^{(j)}} = \bigtr{\cW A\cE^{\top}}, \]
	%which is equal to equation \eqref{eq_proof_same_cov_1} by the definition of $\cW$.
	Hence the proof is complete.

	\iffalse
	Second, notice that
	%\begin{align}
	%	\sum_{j=1}^t \norm{\cW A_j}^2 = \sum_{j=1}^t \bigtr{\cW A_j A_j^{\top} \cW^{\top}} = \bigtr{\cW AA^{\top} \cW^{\top}} = \bigtr{U_X U_X^{\top} \cE A^{\top} (AA^{\top})^{-1} A \cE^{\top}}. \label{eq_proof_same_cov_1}
	%\end{align}
	%We observe that
	\begin{align*}
			\exarg{\cE}{\cE A^{\top} (AA^{\top})^{+} A \cE^{\top}}
		&= \exarg{\cE}{\sum_{j=1}^t \sum_{k=1}^t \varepsilon^{(j)} A_j^{\top} (AA^{\top})^{+} A_k {\varepsilon^{(k)}}^{\top}} \\
		&= \exarg{\cE}{\sum_{j=1}^t \varepsilon^{(j)} A_j^{\top} (AA^{\top})^{+} A_j {\varepsilon^{(j)}}^{\top}} \\
		&= \sigma^2 \cdot r \cdot \id_{n\times n}.
	\end{align*}
	In the above derivation, the second step used the fact that for any $j\neq k$, $\varepsilon^{(j)}$ and $\varepsilon^{(k)}$ are pairwise independent.
	The third step used the fact that $\sum_{j=1}^t A_j^{\top} (AA^{\top})^{+} A_j = \bigtr{\id_{r\times r}} = r$, and $\ex{\varepsilon^{(j)} {\varepsilon^{(j)}}^{\top}} = \sigma^2 \cdot \id_{n\times n}$.
	Therefore, we have that
%	\begin{align*}
%		\inner{\cW A}{\cE} = \sigma^2 \cdot r \cdot \bigtr{U_X U_X^{\top}} = \sigma^2 \cdot r \cdot p,
%	\end{align*}
%because $U_X$ has rank $p$ by Fact \ref{lem_minv}.
	\begin{align*}
		\exarg{\cE}{ \inner{\cW A}{\cE}} = \sigma^2 \cdot r \cdot \bigtr{X(X^{\top} X)^{-1} X^{\top}} =  \sigma^2 \cdot r \cdot \bigtr{ X^{\top}X(X^{\top} X)^{-1}}  = \sigma^2 \cdot r \cdot p.
	\end{align*}
   because $X^\top X$ is a $p\times p$ matrix.	
	%Next, note that
	%\[ \sum_{j=1}^t \inner{\cW A_j}{\varepsilon^{(j)}} = \bigtr{\cW A\cE^{\top}}, \]
	%which is equal to equation \eqref{eq_proof_same_cov_1} by the definition of $\cW$.
	Hence the proof is complete.
\fi

@@

	For equation \eqref{eq_gA_p3}, using Corollary \ref{cor_calE},
	%Lemma \ref{largedeviation} and the fact that all moments of $\varepsilon_i$ exist, 
	we obtain that with high probability,
	\begin{align}
		\abs{\normFro{\cE}^2 - \ex{\normFro{\cE}^2}} \le p^{c} \sigma^2 \| \id_{n\times n}\|_F = p^{c} \cdot \sqrt{n} \cdot \sigma^2 \le p^{-c_{\infty}} \cdot \sigma^2 \cdot n \cdot t, \label{eq_gA_err3}
	\end{align}
	as long as the constant $c$ is taken sufficiently small. 
	For the inner product between $\cW A$ and $\cE$, we have that with high probability,
	\begin{align}
		\bigabs{\inner{\cW A}{\cE} - \exarg{\cE}{\inner{\cW A}{\cE}} } &= \bigabs{\bigtr{\left(\cE^\top U_X U_X^{\top}\cE  - p\sigma^2 \cdot  \id_{t\times t} \right)  U_AU_A^{\top} }} \nonumber  \\
		&\le \bignormFro{U_{A} U_A^{\top}} \cdot \bignormFro{\cE^\top U_X U_X^{\top} \cE - p \sigma^2 \cdot  \id_{t\times t}} \nonumber \\
		&\le \sqrt{r} \cdot (p^{\e}\sigma^2\cdot \|U_X U_X^\top\|_F) \nonumber \\
		&\le p^{1/2+c} \cdot \sigma^2 \cdot \sqrt{r} \nonumber\\
		& \le  p^{-c_{\infty}} \cdot \sigma^2 \cdot n \cdot t .\label{eq_gA_err4}
	\end{align}

%	\begin{align}
%		\bigabs{\inner{\cW A}{\cE} - \ex{\inner{\cW A}{\cE}}} &= \bigabs{\bigtr{U_X U_X^{\top} (\cE U_AU_A^{\top} \cE^{\top} - \sigma^2 \cdot r \id_{n\times n})}} \nonumber  \\
%		&\le \bignormFro{U_{X} U_X^{\top}} \cdot \bignorm{\cE U_A U_A^{\top} \cE^{\top} - \sigma^2 \cdot r \id_{n\times n}} \nonumber \\
%		&\le p^{1/2} \cdot (\sigma^2 r\cdot p^{\e}). \label{eq_gA_err4}
%	\end{align}
	%\begin{align*}
	%	%|\inner{U_X U_X^{\top} \bigbrace{\varepsilon_i W_i^{\top}} (\cW\cW^{\top})^{-1}W_j}{\varepsilon_j}|
	%	\abs{W_i^{\top} (\cW\cW^{\top})^{-1} W_j \cdot \bigbrace{\varepsilon_i^{\top} U_X U_X^{\top} \varepsilon_j}}
	%	&\le |W_i^{\top} (\cW\cW^{\top})^{-1}W_j| \cdot \sigma^2 p^{\e} \bigtr{(U_X U_X^{\top})^2}^{1/2} \le \sigma^2 p^{1/2+\e}
  % &\left|\left(1-\mathop{\mathbb{E}}_{\varepsilon_j} \right)\left[\inner{U_X U_X^{\top} \bigbrace{\varepsilon_j W_j^{\top}} (\cW\cW^{\top})^{-1}W_j}{\varepsilon_j}\right]\right| \le |W_j^{\top} (\cW\cW^{\top})^{-1}W_j| \cdot \sigma^2 p^{\e} \tr \left[(U_X U_X^{\top})^2\right]\le \sigma^2 p^{1/2+\e},\\
  % &\left|\left(1-\mathop{\mathbb{E}}_{\varepsilon_j} \right)\left[\inner{U_X U_X^{\top} \bigbrace{\varepsilon_j W_j^{\top}} (\cW\cW^{\top})^{-1}W_j}{U_X U_X^{\top} \bigbrace{\varepsilon_j W_j^{\top}} (\cW\cW^{\top})^{-1}W_j}\right]\right|\\
  % &\le |W_j^{\top} (\cW\cW^{\top})^{-1}W_j|^2 \cdot \sigma^2 p^{\e} \tr \left[(U_X U_X^{\top})^4\right]\le \sigma^2 p^{1/2+\e},
	%\end{align*}
	%Combining equation \eqref{eq_gA_}, we get that with high probability
	%\begin{align}
	%	\left|g_2(\cal W)-\sigma^2(n\cdot t - p\cdot r)\right|=\left|g_2(\cal W)- \exarg{\varepsilon_1, \dots, \varepsilon_t} {g_2(\cal W)}\right| \lesssim \sigma^2 p^{1/2+\e}. \label{eq_gA_err}
	%\end{align}
	Combining equations \eqref{eq_gA_err1}, \eqref{eq_gA_err2}, \eqref{eq_gA_err3}, and \eqref{eq_gA_err4}, we obtain equation \eqref{eq_gA_err}.
